\chapter{Introduction}
\label{chapter:introduction}

Recent advances in high-throughput sequencing technologies allow a more in-depth analysis of the human genome. 
RNA-Sequencing, in particular, measures the gene activity from vast amounts of genes from specific tissues\cite{kukurba2015rna}.
A major part of research using this data is investigating the inner workings and mechanisms of diseases. 
The analysis of the interplay between genes and their connections to diseases leads to a deeper understanding of the matter, helps to improve diagnosis and results in more effective treatment\cite{soh2007enabling}.

The high dimensionality of the data containing tens of thousands of genes as features coupled with a low number of samples poses challenges to its processing. 
Data analysis requires dimensionality reduction for a feasible application of computational methods like machine learning. 
Reducing the feature space is reasonable because the data is noisy and redundant\cite{ang2016supervised}.
Feature selection (FS) is a technique that addresses these problems. 
Features are ranked by their usefulness allowing the selection of the most valuable features. 

Many computational methods have been employed and developed for FS.
In bioinformatics though, they miss the biological context which is crucial for a deeper understanding because they only consider the raw numerical gene expression values. 
Literature suggests that the integration of biological knowledge has advantages such as the supplement of data analysis, increased performance and the discovery of connections between data and external knowledge \cite{bellazzi2007towards}.

Machine-readable, curated repositories and meta-repositories of biological knowledge in the form of gene-disease-associations, pathway information and ontologies already exist in public knowledge bases\cite{ashburner2000gene,gene2016expansion,kanehisa2000kegg,pinero2015disgenet}. 
This information can be used to enrich the computational analysis of gene expression data, especially in FS. 
Future large-scale studies of gene expression data can reduce computational cost by integrating this public biological information instead of using complex computational FS methods.

Therefore, our goal is to investigate the automatic integration of external knowledge bases in FS, compare the results with proven computational FS approaches and try to combine both methods. 
The contributions of our work are
\begin{itemize}
\item a flexible framework for the comparison of FS methods and integration of external knowledge
\item a novel approach for automatically combining computational FS with prior knowledge
\end{itemize}

The rest of this thesis is organized as follows. 
Chapter \ref{chapter:relatedWork} presents related work. 
Chapter \ref{chapter:conceptApproachSolution} explains the methodology of our framework. 
Chapter \ref{chapter:evaluation} describes the experiments and discusses the results. 
Chapter \ref{chapter:conclusion} concludes our findings.

\subsection{OLD INTRODUCTION:}

According to the WHO, cancer was the second largest cause of death around the world in 2015 being the reason of almost 1 out of 6 deaths.\footnote{\url{http://www.who.int/cancer/en/}} Statistics from 2012 estimate that there were 14.1 million new cancer patients while 8.2 million people died from it. \cite{torre2015global} Moreover, prognosis for the future estimate around 23.6 million new cancer patients per year in 2030. \footnote{\url{http://www.cancerresearchuk.org/health-professional/cancer-statistics/worldwide-cancer}} This shows how severe this disease impacts humankind.

However, cancer is not just one single disease. There are more than 100 different types and subtypes of cancer\footnote{\url{https://www.cancer.gov/types}}. Moreover, treatments and their effectiveness may vary substantially from type to type. Therefore, an accurate diagnosis of cancer type is crucial in order to improve treatment success and reduce costs. 

One approach to advance cancer diagnosis is the utilization of computational methods. In particular, sequencing the transcriptome of samples from patients with various cancer types enables the analysis of the disease for research through computational methods. One aspect of this analysis is to try and classify samples. In recent years, advances in sequencing technology enabled the development of RNA-Sequencing (RNA-Seq), a method that provides higher coverage and greater resolution than previous techniques. Most RNA-Seq experiments feature few samples (in the tens or hundreds) and measure gene expression levels in the order of ten thousand. 

Therefore, classification using machine learning approaches in these scenarios suffers from the curse of dimensionality resulting in feasibility problems, mainly extremely long processing times. To prevent this, feature selection or differential gene expression (DGE) techniques have been utilized and even especially developed for the field of bioinformatics. Their goal is to reduce the set of genes used in machine learning models to a small number that achieves good performance while providing reasonable computation time. Moreover, the selected genes can be seen as associated with the investigated disease.

On the other hand, 	existing knowledge bases like the Comparative Toxicogenomics Database (CTD)\footnote{http://ctdbase.org/} or UniProt\footnote{http://www.uniprot.org/} already contain a lot of information about the connection between diseases and genes. The information about these associations could already be used in the gene selection stage. This would erase the need of employing feature selection.

Therefore, the goal of this master's thesis is to examine whether the utilization of external biological knowledge for gene selection in cancer classification achieves performance comparable to the results of popular feature selection techniques. 

The rest of the thesis is organized as follows. Chapter 2 covers related work and background knowledge. Chapter 3 describes the experiment conducted and lists the results. Those results are evaluated in chapter 4. A discussion follows in chapter 5. Finally, chapter 6 will conclude and discuss possible future work.

+ DGE

From "An integrative gene selection with association analysis for microarray data classification":

This may reduces the accuracy and speed of classification models [5], as conventional classification methods are not designed to cope with high-dimensional datasets [6] and are prone to overfitting [7]. Besides, it is also likely lead to false positives,where gene combinations may be correlated with a target class purely by chance, or a gene is declared to be differentially expressed but actually it is not [8].

\chapter{Motivation}
\label{chapter:Motivation}

Super Motivation.
\chapter{Problem Statement}
\label{chapter:problemStatement}

In this work we consider the problems.
Therefore we want to answer the following research questions in our work.
\begin{enumerate}
\item Q1 ?
\item Q2 ?
\end{enumerate}

For this work the following restrictions apply.
\begin{enumerate}
	\item Restriction 1
	\item Restriction 2
\end{enumerate}

To answer our research questions, the remainder of this work is structured in the following way.
